---
output:
  word_document: default
  html_document: default
---
###Assignment 1
##Chelsea Powell
#BAN 502

```{r load packages}
#install.packages("tidyverse")
library(tidyverse)
```

```{r}
diamonddata = diamonds

glimpse(diamonddata)
```

```{r}
ggplot(diamonddata, aes(x=carat, y=price)) + geom_point()
```

As we can see from the scatterplot above the larger the carat gets the more expensive the diamond can be.

```{r}
ggplot(diamonddata, aes(x=carat, y=price, color=cut)) + geom_point()
```

Based on the scatterplot above it appears that the most expensive diamonds have a premium or ideal cut with a larger carat. The cut of a diamond is important when determining how much the diamond is worth.

```{r}
ggplot(diamonddata, aes(x=carat, y=price, color=cut)) + facet_grid(rows = vars(color)) + geom_point()

```
From the scatterplot above it appears that the larger carat diamonds have a more common color of H, I, or J. The D and E colored diamonds are smaller but are more expensive.

```{r}
library(readr)
Inventory <- read_csv("InventoryData.csv")
View(Inventory)
```

```{r}
InventoryA = filter(Inventory, Supplier == "A")
glimpse(InventoryA)
```
There are 3,695 rows in this new data frame.

```{r}
InventoryA = mutate(InventoryA, OnHandRatio = `On Hand` / `Annual Demand`)

```
This line of code creates a new column that calcuates the on hand ration of a product in our inventory.

```{r}
avg_cost = Inventory %>% group_by(`Item SKU`) %>% summarise(SKUAvgCost = mean(`Cost per Unit ($)`))
```

I found the grouping and filter the most challenging in this assignemnt. The faceting part of one of the questions also confused me. I'm not actually sure if I did it right. I need to dust off the cobwebs and get familiar with R again.








