---
output:
  word_document: default
  html_document: default
---
#BAN 502: Module 4 Assignment 1
##Classification Trees
###Chelsea Powell

```{r}
library(tidyverse)
library(caret)
library(rpart)
library(RColorBrewer)
#install.packages("e1071")
#install.packages("rattle")

library(rattle)
library(e1071)
```

```{r}
library(readr)
parole <- read_csv("parole.csv")
View(parole)

str(parole)
summary(parole)
```

```{r}
parole = parole %>% mutate(male = as.factor(male)) %>% 
  mutate(male = fct_recode(male, "female" = "0", "male" = "1" )) %>%
  mutate(race = as.factor(race)) %>% mutate(race = as.factor(race)) %>%
  mutate(race = fct_recode(race, "white" = "1", "other" = "2" )) %>%
  mutate(state = as.factor(state)) %>%
  mutate(state = fct_recode(state, "Other" = "1", "Kentucky" = "2", "Louisiana" = "3", "Virginia" = "4" )) %>%
  mutate(crime = as.factor(crime)) %>% 
  mutate(crime = fct_recode(crime, "Other" = "1", "Larceny" = "2", "Drug" = "3", "Driving" = "4" )) %>%
  mutate(multiple.offenses = as.factor(multiple.offenses)) %>% 
  mutate(multiple.offenses = fct_recode(multiple.offenses, "Single Offense" = "0", "Multiple Offenses" = "1" )) %>%
  mutate(violator = as.factor(violator)) %>% 
  mutate(violator = fct_recode(violator, "completed" = "0", "violated" = "1" ))
str(parole)
```

```{r}
set.seed(12345)
train.rows = createDataPartition(y=parole$violator, p=0.7, list = FALSE)
train = parole[train.rows,]
test = parole[-train.rows,]
```

```{r}
tree1 = rpart(violator  ~., train, method="class")
fancyRpartPlot(tree1)
```
Based on the classification tree I would say that he has a 7% chance of completing his parole. To determine this I went through the tree and evluated him based on the levels. He is from Louisiana so we go to the right in our decision tree for the state=other, kentucky, virginia. He is less than 43 years old so I go the left for yes and his time served is greater than 2.6 so I go left again and arrive to my 7% chance of completing his parole conclusion.

```{r}
printcp(tree1)
plotcp(tree1)
```
The CP level of .054545 shold be chosen because it has the lowest relative error value.

```{r}
tree2 = prune(tree1,cp= tree1$cptable[which.min(tree1$cptable[,"xerror"]),"CP"])
printcp(tree2)
```
Age is the majority class of the training set.

```{r}
treepred = predict(tree1, train, type = "class")
head(treepred)
```

```{r}
confusionMatrix(treepred,train$violator,positive="completed")
```


```{r}
treepred_test = predict(tree1, newdata = test, type="class")
confusionMatrix(treepred_test,test$violator,positive="completed")
```
This model is a little less accurate than our testing model. If we were to assume that everyone completed their parole we would be right 88.6% of the time.

```{r}
library(readr)
blood <- read_csv("blood.csv")
View(blood)
```
```{r}
blood = blood %>% mutate(DonatedMarch = as.factor(DonatedMarch)) %>% 
  mutate(DonatedMarch = fct_recode(DonatedMarch, "No" = "0", "Yes" = "1" ))
```

```{r}
set.seed(1234)
train.rows = createDataPartition(y=parole$violator, p=0.7, list = FALSE)
train2 = blood[train.rows,]
test2 = blood[-train.rows,]
```

```{r}
tree3 = rpart(DonatedMarch  ~., train2, method="class")
fancyRpartPlot(tree3)
```

```{r}
printcp(tree3)
plotcp(tree3)
```

```{r}
tree4 = prune(tree3,cp= tree3$cptable[which.min(tree3$cptable[,"xerror"]),"CP"])
printcp(tree4)
```

```{r}
treepred_train1 = predict(tree3, newdata = train2, type="class")
confusionMatrix(treepred_train1,train2$DonatedMarch,positive="Yes")
```
The training model is 79.92% accurate. If we assumed that everyone gave then we would be right 75.9% of the time.
```{r}
treepred_test1 = predict(tree3, newdata = test2, type="class")
confusionMatrix(treepred_test1,test2$DonatedMarch,positive="Yes")
```
The testing model is a little less accurate than our training model. The testing model is 79.64% accurate.


