---
output:
  word_document: default
  html_document: default
---
#BAN 502: Module 3 Assignment 1
##Chelsea Powell
###Model Validation

```{r}
library(tidyverse)
library(MASS)
#install.packages("caret")
library(caret)
```

```{r}
library(readr)
bike <- read_csv("hour.csv")

bike = bike %>% mutate(season = as_factor(as.character(season))) %>%
mutate(season = fct_recode(season,
"Spring" = "1",
"Summer" = "2",
"Fall" = "3",
"Winter" = "4")) %>%
mutate(yr = as_factor(as.character(yr))) %>%
mutate(mnth = as_factor(as.character(mnth))) %>%
mutate(hr = as_factor(as.character(hr)))%>% 
mutate(holiday = as_factor(as.character(holiday))) %>%
mutate(holiday = fct_recode(holiday,
"Not Holiday" = "0",
"Holiday" = "1")) %>%
mutate(workingday = as_factor(as.character(workingday))) %>%
  mutate(workingday = fct_recode(workingday,
"NotWorkingDay" = "0",
"WorkingDay" = "1")) %>%
mutate(weathersit = as_factor(as.character(weathersit))) %>%
mutate(weathersit = fct_recode(weathersit,
"NoPrecip" = "1",
"Misty" = "2",
"LightPrecip" = "3",
"HeavyPrecip" = "4")) %>%
mutate(weekday = as_factor(as.character(weekday))) %>%
mutate(weekday = fct_recode(weekday,
"Monday" = "1",
"Tuesday" = "2",
"Wednesday" = "3",
"Thursday" = "4",
"Friday" = "5",
"Saturday" = "6",
"Sunday" = "0"))
```

```{r}
set.seed(1234)
train.rows = createDataPartition(y = bike$count, p=0.7, list = FALSE) #70% in training
train = bike[train.rows,] 
test = bike[-train.rows,]

```
There are 5,212 rows in the testing data and 12167 rows in the training data.

```{r}
mod1 = lm(count ~ season+mnth+hr+holiday+weekday+temp+weathersit, bike) #create linear regression model
summary(mod1)
```
The adjusted R squared for this model is 0.62 which is pretty good. The p-value is signigicant because it is less than 0.05. I would say this is a good model to use to train and test our predictions.

```{r}
predict_train = predict(mod1, newdata = train)
head(predict_train)
```
The predictions in the header training data range range from -60.46 to 11.54. There are 12,167 values in our predicted training data.

```{r}
predict_test = predict(mod1, newdata = test)
head(predict_test)
```
The predictions in the training data range from -16.19 to 215.76. There are 5,212 values in our predicted testing data.

```{r}
SSE = sum((test$count - predict_test)^2) #sum of squared residuals from model
SST = sum((test$count - mean(test$count))^2) #sum of squared residuals from a "naive" model
1 - SSE/SST
```
The manually calculated R squared value is slightly different from the k fold cross validations r squared above. The r-squared value from the test and training data is .0053 higher than the cross validation r-squared value.
















