---
output:
  word_document: default
  html_document: default
---
#BAN 502: Module 4 Assignment 2
##Random Forests
###Chelsea Powell


```{r}
#install.packages("ranger")
library(tidyverse)
library(caret)
library(ranger)
```

```{r}
library(readr)
blood <- read_csv("Blood.csv")

blood = blood %>% mutate(DonatedMarch = as.factor(DonatedMarch)) %>% 
  mutate(DonatedMarch = fct_recode(DonatedMarch, "No" = "0", "Yes" = "1" ))

str(blood)
summary(blood)
```

```{r}
set.seed(1234) 
train.rows = createDataPartition(y = blood$DonatedMarch, p=0.7, list = FALSE) #70% in training
train = blood[train.rows,] 
test = blood[-train.rows,]
```

```{r}
fit_control = trainControl(method = "cv",  
                           number = 10) 

set.seed(123)  
rf_fit = train(DonatedMarch ~.,    
                 data = train,   
                 method = "ranger",  
                 importance = "permutation", 
                 num.trees = 100,
                 trControl = fit_control)
```

```{r}
varImp(rf_fit)
rf_fit
```
TotalDonations is the most important variable in this dataset for this model. Mnts_Since_First is the least important variable.

```{r}
predRF_test = predict(rf_fit, train)
head(predRF_test)
```

```{r}
confusionMatrix(predRF_test, train$DonatedMarch, positive = "Yes")
```
The accuracy of this model is 89.5% which is better than the no information rate of 76.15%. The sensivity of this model is 0.6 and the specificity of this model is 0.9875.

```{r}
predRF_test = predict(rf_fit, newdata = test)
```

```{r}
confusionMatrix(predRF_test, test$DonatedMarch, positive = "Yes")
```
The accuracy went down on our testing set at 78.12%. Compared to the no information rate of 76.34 our model is better but not by a ton.
