---
output:
  word_document: default
  html_document: default
---
#BAN 502: Module 2 Assignment 1
##Chelsea Powell
###Simple Linear Regression and Correlation


```{r setup}
library(tidyverse)
#install.packages("GGally")
library(GGally)

air = airquality
glimpse(air)
```
a. This dataset describes New York air quality in New York from May to September 1973.
b. There are 6 varialbes and 153 observations.
c. There is missing data in the Ozone and Solar.R attributes.
d. Ozone is most likely the response (Y) variable.

```{r delete missing values}
air2 = air %>% filter(!is.na(Ozone)) %>% filter(!is.na(Solar.R))
glimpse(air2)
```
After removing any rows with missing data there are 111 observations and 6 variables. Fourty two rows were deleted from this data set.

```{r}
ggpairs(air2)
ggcorr(air2, label = TRUE)
```
a. Temperature is most strongly correlated to the Ozone variable.
b. Day is least strongly correlated to the Ozone variable.

```{r}
ggplot(air2, aes(x=Temp, y=Ozone)) + geom_point()
```
It appears that the Ozone and Temperature variables have a positive correlation. As the temperature variable increase so does the ozone.

```{r}
model1 = lm(Ozone ~ Temp, data=air2)
summary(model1)

confint(model1)
```
a. The R squared value is ok at 48.8%. The predictor varialbe of temperature is signifigant becuase the p-value is less than 0.05.
b. The slope coefficient will likely fall in between -184.82 and -110.47.

```{r}
ggplot(air2, aes(x=Temp, y=Ozone)) + geom_point() + geom_smooth(method="lm", se=FALSE)
```

```{r}
testdata = data.frame(Temp = 80)
predict(model1, newdata = testdata, interval = "predict")
```

```{r}
ggplot(air2, aes(x=Day, y=Ozone)) + geom_point()
```
By just looking at the chart above it looks like there is no relationship between the two variables day and ozone.

```{r}
model2 = lm(Ozone ~ Day, data=air2)
summary(model2)

confint(model2)

```
a. The R squared value is very poor. It is very small value less than 1%. The p-value is not significant becuase it is .96 which is much larger than 0.05.
b. The slope coefficient falls between 29.25 and 55.58.

```{r}
ggplot(air2, aes(x=Day, y=Ozone)) + geom_point() + geom_smooth(method="lm", se = FALSE)
```

